---
title: "rSNPdata"
author: "Karim MANÉ"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{snpdata_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

---
date: `r format(Sys.Date(), "%d/%m/%Y")`
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,comment = "#>")
require("rSNPdata")
require("data.table")
require("dplyr")
require("SeqArray")
require("moimix")
require("statip")
require("parallel")
vcf="/Users/km28/Documents/Karim/Packages/Input_Data.vcf.gz"
metadata="/Users/km28/Documents/Karim/Packages/SampleMetadata.txt"
outDir="/Users/km28/Documents/Karim/Packages/Test"
```

## Install   
``` {r eval=FALSE}
library(devtools)
devtools::install_github("FellouMada/rSNPdata", build_vignettes = TRUE)
library(rSNPdata)
```

# FUNCTIONS     

## get SNPdata object      
the functions in this package require a `SNPdata` object. This can be generated using the `get_snpdata` function with the following arguments:    
1. the input VCF file     
2. the file with the samples metadata     
3. the path to the folder where the output files will be stored     
type `?get_snpdata` for more details      
```{r eval=FALSE}
snpdata=get_snpdata(vcf.file = vcf, meta.file = metadata, output.dir = outDir)
```

## print the SNPdata object     
```{r, eval=FALSE}
print(snpdata)
```

## calculate the SNPs minor allele frequency     
the `compute_MAF` requires a SNPdata object, the name of the genotype table from which the MAF should be calculated (either "GT" for the raw genotypes, "Phased" for phased genotypes, "Phased_Imputed" for phased and imputed genotypes).    
include.het determines whether to account for heterozygous sites in MAF calculation.   
type `?compute_MAF` for more details     
```{r, eval=FALSE}
snpdata = compute_MAF(snpdata, include.het=FALSE, mat.name="GT")
head(snpdata$details)
```

## calculate the within-host genetics diversity index (Fws)   
the `calculate_Fws` function requires a SNPdata object   
type `?compute_MAF` for more details     
```{r, eval=FALSE}
snpdata =  calculate_Fws(snpdata)
head(snpdata$meta)
```

## filter the SNPdata object     
use the `filter_snps_samples` function to remove the SNPs and samples that do not satisfy the user criteria.    
```{r eval=FALSE}
snpdata = filter_snps_samples(snpdata, min.qual=10, max.missing.sites=0.2, max.missing.samples=0.2, maf.cutoff=0.01)
print(snpdata)
```

## mixed genotypes phasing     
use the `phase_mixed_genotypes` function to phase the mixed genotypes    
```{r eval=FALSE}
snpdata = phase_mixed_genotypes(snpdata)
head(snpdata$Phased[,1:5])
```

## impute the missing data   
use the `impute_missing_genotypes` function to impute the missing genotypes. the `genotype` argument determines which genotype table should be used to impute the missing data ("GT" to impute from raw data, "Phased" to impute from the phased data).    
```{r eval=FALSE}
snpdata = impute_missing_genotypes(snpdata, genotype="GT")
head(snpdata$Imputed[,1:5])
```

## select data for a given chromosome     
to get the data from chromosome 7 ("Pf3D7_07_v3"), use the `select_chrom`    
```{r eval=FALSE}
chrom7_snpdata = select_chrom(snpdata, chrom="Pf3D7_07_v3")
print(chrom7_snpdata)
```

## drop a set of SNPs from the data       
if you want to remove some loci from the data, use the `drop_snps` function and provide it with a data frame with 2 columns, "Chrom" and "Pos", representing the genomic coordinates of the loci to be dropped.     
```{r eval=FALSE}
snp.to.be.dropped = chrom7_snpdata$details %>% select(Chrom,Pos)
idx = sample(1:nrow(snp.to.be.dropped),100, replace = FALSE)
snp.to.be.dropped = snp.to.be.dropped[idx,]
reduced_snpdata = drop_snps(chrom7_snpdata, snp.to.be.dropped)
print(reduced_snpdata)
```

## drop a set of samples from the data       
if you want to remove some samples from the data, use the `drop_samples` function and provide it with a vector of samples to be dropped.     
```{r eval=FALSE}
idx = sample(1:nrow(reduced_snpdata$meta),10, replace = FALSE)
samples.to.be.dropped = reduced_snpdata$meta$sample[idx]
reduced_snpdata = drop_samples(reduced_snpdata, samples.to.be.dropped)
print(reduced_snpdata)
```




