---
title: "rSNPdata"
author: "Karim MANÉ"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{snpdata_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

---
date: `r format(Sys.Date(), "%d/%m/%Y")`
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,comment = "#>")
require("data.table")
require("dplyr")
require("SeqArray")
require("moimix")
require("statip")
require("parallel")
require("isoRelate")
require("Rfast")
require("utils")
vcf="/Users/km28/Documents/Karim/Packages/Input_Data.vcf.gz"
metadata="/Users/km28/Documents/Karim/Packages/SampleMetadata.txt"
outDir="/Users/km28/Documents/Karim/Packages/Test"
```

## Install   
``` {r eval=FALSE}
library(devtools)
devtools::install_github("FellouMada/rSNPdata", build_vignettes = TRUE)
library(rSNPdata)
```


## get SNPdata object      
the functions in this package require a `SNPdata` object. This can be generated using the `get_snpdata` function with the following arguments:    
1. the input VCF file     
2. the file with the samples metadata     
3. the path to the folder where the output files will be stored     
type `?get_snpdata` for more details      
```{r eval=FALSE}
snpdata=get_snpdata(vcf.file = vcf, meta.file = metadata, output.dir = outDir)
```

## print the SNPdata object     
```{r, eval=FALSE}
print(snpdata)
```

## calculate the SNPs minor allele frequency     
the `compute_MAF` requires a SNPdata object, the name of the genotype table from which the MAF should be calculated (either "GT" for the raw genotypes, "Phased" for phased genotypes, "Phased_Imputed" for phased and imputed genotypes).    
include.het determines whether to account for heterozygous sites in MAF calculation.   
type `?compute_MAF` for more details     
```{r, eval=FALSE}
snpdata = compute_MAF(snpdata, include.het=FALSE, mat.name="GT")
head(snpdata$details)
```

## filter the SNPdata object     
use the `filter_snps_samples` function to remove the SNPs and samples that do not satisfy the user criteria.    
```{r eval=FALSE}
snpdata = filter_snps_samples(snpdata, min.qual=10, max.missing.sites=0.2, max.missing.samples=0.2, maf.cutoff=0.01)
print(snpdata)
hist(snpdata$details$MAF, 100)
grid()
```

## calculate the within-host genetics diversity index (Fws)   
the `calculate_Fws` function requires a SNPdata object   
type `?calculate_Fws` for more details     
```{r, eval=FALSE}
snpdata =  calculate_Fws(snpdata)
head(snpdata$meta)
```

## mixed genotypes phasing     
use the `phase_mixed_genotypes` function to phase the mixed genotypes.    
```{r eval=FALSE}
snpdata = phase_mixed_genotypes(snpdata, nsim=100)
head(snpdata$Phased[,1:5])
```

## impute the missing data   
use the `impute_missing_genotypes` function to impute the missing genotypes. the `genotype` argument determines which genotype table should be used to impute the missing data ("GT" to impute from raw data, "Phased" to impute from the phased data).    
```{r eval=FALSE}
snpdata = impute_missing_genotypes(snpdata, genotype="GT", nsim=100)
head(snpdata$Imputed[,1:5])
```

## select data for a given chromosome     
to get the data from chromosome 7 ("Pf3D7_07_v3"), use the `select_chrom`    
```{r eval=FALSE}
chrom7_snpdata = select_chrom(snpdata, chrom="Pf3D7_07_v3")
print(chrom7_snpdata)
```

## drop a set of SNPs from the data       
if you want to remove some loci from the data, use the `drop_snps` function and provide it with a data frame with 2 columns, "Chrom" and "Pos", representing the genomic coordinates of the loci to be dropped or the chrom, start and end position of the loci to be dropped    
```{r eval=FALSE}
snp.to.be.dropped = chrom7_snpdata$details %>% select(Chrom,Pos)
idx = sample(1:nrow(snp.to.be.dropped),100, replace = FALSE)
snp.to.be.dropped = snp.to.be.dropped[idx,]
reduced_snpdata = drop_snps(chrom7_snpdata, snp.to.be.dropped)
print(reduced_snpdata)
reduced_snpdata = drop_snps(chrom7_snpdata, chrom="Pf3D7_10_v3", start=100, end=500)
print(reduced_snpdata)
```

## drop a set of samples from the data       
if you want to remove some samples from the data, use the `drop_samples` function and provide it with a vector of samples to be dropped.     
```{r eval=FALSE}
idx = sample(1:nrow(reduced_snpdata$meta),10, replace = FALSE)
samples.to.be.dropped = reduced_snpdata$meta$sample[idx]
reduced_snpdata = drop_samples(reduced_snpdata, samples.to.be.dropped)
print(reduced_snpdata)
```


## Fst
In this package, Weir & Cockerham's Fst is calculated using the `vcflib` tools.   
The from option specifies the name of the column in the metadata table with the groups between which Fst should be calculated. 
the `calculate_wcFst` returns a `SNPdata` object with an additionnal field named `Fst`. This is a list that contains data frames representing each the Fst results between a given pair-wise population.   
```{r eval=FALSE}
snpdata=get_snpdata(vcf.file = vcf, meta.file = metadata, output.dir = outDir)
snpdata = compute_MAF(snpdata, include.het=FALSE, mat.name="GT")
snpdata = filter_snps_samples(snpdata, min.qual=10, max.missing.sites=0.2, max.missing.samples=0.2, maf.cutoff=0.01)
snpdata = calculate_wcFst(snpdata, from="Country", groups=c("Senegal","Gambia"))
names(snpdata$Fst)
head(snpdata$Fst$Senegal_vs_Gambia)
```

## IBS   
The identity by state (IBS) represents the proportion of similar loci between 2 pairs of isolates. The `calculate_IBS` function returns the dissimilarity matrix i.e. (1-IBS). It returns a `SNPdata` object with an additionnal field named `IBS`.    
```{r eval=FALSE}
snpdata = calculate_IBS(snpdata, mat.name="GT")
```

## LD   
The linkage between pairs of SNPs is calculated using the `vcftools` program.   
```{r eval=FALSE}
## calculate LD between pairs of SNPs across chromosome 4 and 5
snpdata = calculate_LD(snpdata, min.r2=0.2, inter.chrom=FALSE, chroms=c("Pf3D7_04_v3","Pf3D7_05_v3"))

## calculate LD between pairs of SNPs across all chromosomes
snpdata = calculate_LD(snpdata, min.r2=0.2, inter.chrom=FALSE, chroms=NULL)
```

## iR
```{r eval=FALSE}
groups = c("DongoroBa","Chogen")
idx = which(snpdata$meta$Location %in% groups)
samples.to.be.dropped = snpdata$meta$sample[-idx]
reduced_snpdata = drop_samples(snpdata, samples.to.be.dropped)
print(reduced_snpdata)
reduced_snpdata =  calculate_Fws(reduced_snpdata)
res.iR = calculate_iR(reduced_snpdata, mat.name="Phased", family="Location", number.cores=4)
dim(res.iR$iR$Chogen_vs_DongoroBa)
```

## relatedness
relatedness between pairs of isolates from a given pairs of population. The relatedness is calculated using the `GT` table. The value of the `from` option is the name of the column in the metadata data frame that represents the sample's population of origin. To remove regions of the genome under selection, use the `sweepRegions` option. You can specify a vector of populations to compute the relatedness for as a value for the groups option.  
```{r eval=FALSE}
## calculate relatedness between pairs of isolates from Chogen and DongoroBa
snpdata = phase_mixed_genotypes(snpdata, nsim=5)
snpdata = impute_missing_genotypes(snpdata, genotype="GT", nsim=10)
snpdata = calculate_relatedness(snpdata, mat.name="GT", from="Location", sweepRegions=NULL, groups=c("Chogen","DongoroBa"))
relatedness = snpdata$relatedness$df
relatedness.matrix = snpdata$relatedness$matrix

## calculate relatedness between pairs of isolates from all locations
snpdata = calculate_relatedness(snpdata, mat.name="GT", family="Location", sweepRegions=NULL, groups=NULL)

## calculate relatedness between pairs of isolates from all locations after excluding the selective sweep regions
Chrom = c("Pf3D7_04_v3","Pf3D7_05_v3","Pf3D7_07_v3")
Start = c(723088, 932890, 378222)
End = c(774914, 987149, 431317)
selection.region = data.frame(cbind(Chrom, Start, End), stringsAsFactors = FALSE)
snpdata = calculate_relatedness(snpdata, mat.name="GT", family="Location", sweepRegions=selection.region, groups=NULL)


```

